---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# FlyingR

<!-- badges: start -->
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/BMasinde/FlyingR?branch=dev-0.2.1&svg=true)](https://ci.appveyor.com/project/BMasinde/FlyingR)
[![codecov](https://codecov.io/gh/BMasinde/FlyingR/branch/dev-0.2.1/graph/badge.svg)](https://codecov.io/gh/BMasinde/FlyingR)
<!-- badges: end -->

The package provides methods for predicting flight range of birds based on their
physiological characteristics. This is an R implementation of Flight program
provided by Pennycuick

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BMasinde/FlyingR")
```
## Examples

###  Time-Marching computation 
```{r example}
library(FlyingR)
## basic example code

## birds comes with the package
data("birds")

simulation <- migrate(data = birds,  method = "cmm", settings = list(airDensity = 0.905))

simulation$range
```

The function also returns the mechanical and chemical power during the simulation

### Range estimation based on ODE
This function estimates the range based on Pennycuick (1975) Mechanics of Flight where Breguet set of equations are used.
```{r}
## when estimating range of a single bird
birds_range <- flysim(data = birds,  settings = list(airDensity = 0.905))

birds_range$range
```

## The data
*birds* definitions pulled from Flight program in-built datasets and fat mass
randomly generated where initially zero. In addition, by default muscle mass was
derived as 0.17 fraction of the all-up mass. User's data should have columns named
appropriately. The package looks for columns named *id, name or species.name*,
*bodymass or allupmass*, *wingspan, ws*, *wingarea*, *ordo, order* (which is a
factored column with levels 1 or 2 passerines and non-passerines respectively)
*fatmass, fat.mass, fat_mass* and lastly *muscle_mass*.

```{r }
birds
```





