---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# FlyingR

<!-- badges: start -->
[![R-CMD-check](https://github.com/BMasinde/FlyingR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BMasinde/FlyingR/actions/workflows/R-CMD-check.yaml)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/BMasinde/FlyingR?branch=master&svg=true)](https://ci.appveyor.com/project/BMasinde/FlyingR)
[![codecov](https://codecov.io/gh/BMasinde/FlyingR/branch/master/graph/badge.svg)](https://app.codecov.io/gh/BMasinde/FlyingR)
<!-- badges: end -->

The package provides methods for predicting flight range of birds based on their
physiological characteristics. This is an R implementation of Flight program
provided by Pennycuick.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BMasinde/FlyingR")
```


##  Time-Marching computation
In brief, the time-marching computation computes the flight range of birds in 
short time intervals (6-minutes). Within this time period chemical and mechanical 
powers are held constant during which fat and protein are consumed to sustain 
flight. Because of the reduction in weight, the chemical and mechanical powers are
recalculated for the next time period and the distance flown is incremented to 
previous time-interval's distance. And this process iterates until fat mass is 
depleted giving the maximum possible range under the time-marching computation. 


## 
Birds migrate long distances between breeding and wintering grounds. In the 
majority of cases, it is not possible to cover these distances at a go, and 
therefore, stops are required to rest, feed, and restore both fat and protein 
Lindström and Piersma 1993; Klaassen 1996; Kaiser 1999; Schwilch et al. 2002; Pennycuick 2008).
However, research on stopovers (e.g., Lindström 1991; Lindström and Piersma 1993)  
focuses on fuel (fat mass) restoration. As such, the stopover mass calculator 
used in the package FlyingR, is based on maximum fuel deposition rates (FDR), 
given by Lindström (1991), separately for passerines and non-passerines (Equations \ref{fdr passerines}, \ref{fdr nonpasserines}, and \ref{fatmassgained}). This is not an optimal solution as it 
does not account for protein restoration. In addition, there are reservations about using lean body mass; body mass minus fat mass (Pennycuick 2008).

$$
    FDR_{max; \ passerines} = 2.22 \times lean \ body \ mass^{-0.27}
    \label{fdr passerines}
$$

$$
    FDR_{max; \ non-passerines} = 2.80 \times lean \ body \ mass^{-0.27}
    \label{fdr nonpasserines}
$$

$$
    fat \ mass \ gained = lean \ body \ mass \times \frac{FDR_{max}}{100} \times duration
    \label{fatmassgained}
$$

## Examples

### Examples 1: Garden Warbler and Lesser Whitethroat
These two passerine species are trans-Saharan migrants. Lesser whitethroats winter 
in the Sahel zone of eastern and north-eastern Africa \citep{pearson1992migration}, 
while garden warblers spend winter further, i.e., in southern Africa 
\citep[][]{shirihai2001sylvia}. Data on the Garden Warbler and Lesser 
Whitethroat are obtained from various bird ringing stations situated along their 
SE European flyway leading from Europe towards African winter quarters. This is 
split into four regions: Southern Baltic, Eastern Mediterranean,
Northern Mediterranean and North Eastern Africa. These data include the 1-st year 
individuals (immatures): 1,044 garden warblers and 848 lesser whitethroats, and 
span the years 2000 to 2006. The sample contains the following variables: station 
code, ring number, age, body mass, fat score, wing and tail measurements. The fat 
score is used to derive fat fraction and subsequently fat mass by the procedure 
described by \citet{ozarowska2015contrasting}. Body mass and fat mass are converted 
from grams to kilograms. The wing measurements do not equal the required wingspan 
as these are measured differently (i.e., the wingspan is measured from tip to tip 
of wings, particularly the primary feathers, in metres). Other missing variables 
are the wing area and muscle mass. \citet{bruderer2001flight} provide estimates 
for the wingspan and wing area for several species. For the Garden Warbler, the 
wingspan and wing area are 0.223 (m) and 0.0093 ($m^2$), respectively. While for 
the Lesser Whitethroat, these are 0.185 (m) and 0.0073 ($m^2$). Muscle mass 
estimates are obtained by using the muscle fraction of 0.17 as recommended by 
\citet{penny08}. Figure \ref{results_range} presents the flight range distribution 
of these two species in kilometres for each region separately. 

```{r example}
library(FlyingR)
## basic example code

## birds comes with the package
data("birds")

simulation <- migrate(data = birds,  method = "cmm", settings = list(airDensity = 0.905))

simulation$range
```

The function also returns the mechanical and chemical power during the simulation

### Range estimation based on ODE
This function estimates the range based on Pennycuick (1975) Mechanics of Flight where Breguet set of equations are used.
```{r}
## when estimating range of a single bird
birds_range <- flysim(data = birds,  settings = list(airDensity = 0.905))

birds_range$range
```

## The data
*birds* definitions pulled from Flight program in-built datasets and fat mass
randomly generated where initially zero. In addition, by default muscle mass was
derived as 0.17 fraction of the all-up mass. User's data should have columns named
appropriately. The package looks for columns named *id, name or species.name*,
*bodymass or allupmass*, *wingspan, ws*, *wingarea*, *ordo, order* (which is a
factored column with levels 1 or 2 passerines and non-passerines respectively)
*fatmass, fat.mass, fat_mass* and lastly *muscle_mass*.

```{r }
birds
```


## References
Lindström Å, Piersma T (1993). “Mass changes in migrating birds: the evidence for fat
and protein storage re-examined.” Ibis, 135(1), 70–78. doi:10.1111/j.1474-919X.1993.
tb02811.x

Klaassen M (1996). “Metabolic constraints on long-distance migration in birds.” Journal of
Experimental Biology, 199(1), 57–64. doi:10.1242/jeb.199.1.57.

Kaiser A (1999). “Stopover strategies in birds: a review of methods for estimating stopover
length.” Bird Study, 46(sup1), S299–S308. doi:10.1080/00063659909477257.

Schwilch R, Grattarola A, Spina F, Jenni L (2002). “Protein loss during long-distance migra-
tory flight in passerine birds: adaptation and constraint.” Journal of Experimental Biology,
205(5), 687–695. doi:10.1242/jeb.205.5.687.

Pennycuick CJ (2008). MODELLING THE FLYING BIRD, volume 5. First edition edition.
Elsevier, New Jersey.

Lindström Å, Piersma T (1993). “Mass changes in migrating birds: the evidence for fat
and protein storage re-examined.” Ibis, 135(1), 70–78. doi:10.1111/j.1474-919X.1993.
tb02811.x.

