---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# flying

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/BMasinde/flight.svg?branch=master)](https://travis-ci.org/BMasinde/flight)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/BMasinde/flight?branch=master&svg=true)](https://ci.appveyor.com/project/BMasinde/flight)
[![Coveralls test coverage](https://coveralls.io/repos/github/BMasinde/flight/badge.svg)](https://coveralls.io/r/BMasinde/flight?branch=master)
<!-- badges: end -->

The package provides methods for predicting flight range of birds based on their
physiological characteristics. This is an R implementation of Flight program
provided by Pennycuick.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BMasinde/flight")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(flying)
## basic example code

## birds comes with the package
data("birds")

birds_range <- flysim(data = birds)

birds_range$range
```

```{r}
## when estimating range of a single bird
Garden_Warbler <- list("name" = "Garden Warbler",
                       "body_mass" = 0.022,
                       "fat_mass" = 0.00660,
                       "wing_span" = 0.240,
                       "wing_area" = 0.0110,
                       "order" = as.factor(1)
)

G_warbler <- flysim(data = Garden_Warbler, ctrl = list(airDensity = 0.9093))

G_warbler$range
```



The function *flysim* also outputs: constants used, fuel, Minimum power speed
*Vmp*, Maximum range speed *Vmr* (still needs looking into), and lastly the data).

## The data
*birds* definitions pulled from Flight program in-built datasets and fat mass g
randomly generated where initially zero. Users's data should have columns named
appropriately. The package looks for columns named *id, name or species.name*,
*bodymass or allupmass*, *wingspan, ws*, *wingarea*, *ordo, order* (which is a
factored column with levles 1 or 2 passerines and non-passerines respectively.
Lastly *fatmass, fat.mass, fat_mass*.

```{r }
birds
```





