---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
devtools::load_all()
```
# FlyingR

<!-- badges: start -->
[![R-CMD-check](https://github.com/BMasinde/FlyingR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BMasinde/FlyingR/actions/workflows/R-CMD-check.yaml)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/BMasinde/FlyingR?branch=master&svg=true)](https://ci.appveyor.com/project/BMasinde/FlyingR)
[![codecov](https://codecov.io/gh/BMasinde/FlyingR/branch/master/graph/badge.svg)](https://app.codecov.io/gh/BMasinde/FlyingR)
<!-- badges: end -->

The package provides methods for predicting flight range of birds based on their
physiological characteristics. This is an R implementation of Flight program
provided by Pennycuick.

## Installation
Install from CRAN with:

``` r
# use install.packages() function
# or the graphical user interface in RStudio

install.packages("FlyingR)
```

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BMasinde/FlyingR")
```


##  Time-Marching computation
In brief, the time-marching computation computes the flight range of birds in 
short time intervals (6-minutes). Within this time period chemical and mechanical 
powers are held constant during which fat and protein are consumed to sustain 
flight. Because of the reduction in weight, the chemical and mechanical powers are
recalculated for the next time period and the distance flown is incremented to 
previous time-interval's distance. And this process iterates until fat mass is 
depleted giving the maximum possible range under the time-marching computation. 


## Stop-over mass calculator
Birds migrate long distances between breeding and wintering grounds. In the 
majority of cases, it is not possible to cover these distances at a go, and 
therefore, stops are required to rest, feed, and restore both fat and protein 
Lindström and Piersma 1993; Klaassen 1996; Kaiser 1999; Schwilch et al. 2002; Pennycuick 2008).
However, research on stopovers (e.g., Lindström 1991; Lindström and Piersma 1993)  
focuses on fuel (fat mass) restoration. As such, the stopover mass calculator 
used in the package FlyingR, is based on maximum fuel deposition rates (FDR), 
given by Lindström (1991), separately for passerines and non-passerines (Equations \ref{fdr passerines}, \ref{fdr nonpasserines}, and \ref{fatmassgained}). This is not an optimal solution as it 
does not account for protein restoration. In addition, there are reservations about using lean body mass; body mass minus fat mass (Pennycuick 2008).

$$
    FDR_{max; \ passerines} = 2.22 \times lean \ body \ mass^{-0.27}
    \label{fdr passerines}
$$

$$
    FDR_{max; \ non-passerines} = 2.80 \times lean \ body \ mass^{-0.27}
    \label{fdr nonpasserines}
$$

$$
    fat \ mass \ gained = lean \ body \ mass \times \frac{FDR_{max}}{100} \times duration
    \label{fatmassgained}
$$

## Expected variables in datasets
The package looks for columns named *id, name or species.name*,
*bodymass or allupmass*, *wingspan, ws*, *wingarea*, *ordo, order* (which is a
factored column with levels 1 or 2 passerines and non-passerines respectively)
*fatmass, fat.mass, fat_mass* and lastly *muscle_mass*.

## Examples

### Examples 1: Garden Warbler and Lesser Whitethroat
These two passerine species are trans-Saharan migrants. Lesser whitethroats winter 
in the Sahel zone of eastern and north-eastern Africa (Pearson and Lack 1992), 
while garden warblers spend winter further, i.e., in southern Africa 
(Shirihai, Gargallo, and Helbig 2001). Data on the Garden Warbler and Lesser 
Whitethroat are obtained from various bird ringing stations situated along their 
SE European flyway leading from Europe towards African winter quarters. This is 
split into four regions: Southern Baltic, Eastern Mediterranean,
Northern Mediterranean and North Eastern Africa. These data include the 1-st year 
individuals (immatures): 1,044 garden warblers and 848 lesser whitethroats, and 
span the years 2000 to 2006. The sample contains the following variables: station 
code, ring number, age, body mass, fat score, wing and tail measurements. The fat 
score is used to derive fat fraction and subsequently fat mass by the procedure 
described by Ożarowska (2015). Body mass and fat mass are converted 
from grams to kilograms. The wing measurements do not equal the required wingspan 
as these are measured differently (i.e., the wingspan is measured from tip to tip 
of wings, particularly the primary feathers, in metres). Other missing variables 
are the wing area and muscle mass. Bruderer and Boldt (2001) provide estimates 
for the wingspan and wing area for several species. For the Garden Warbler, the 
wingspan and wing area are 0.223 (m) and 0.0093 ($m^2$), respectively. While for 
the Lesser Whitethroat, these are 0.185 (m) and 0.0073 ($m^2$). Muscle mass 
estimates are obtained by using the muscle fraction of 0.17 as recommended by 
 Pennycuick (2008). Figure \ref{results_range} presents the flight range distribution 
of these two species in kilometres for each region separately. 



```{r}
# ------------------------------------------------------------------------------
# A generic function 
# migration function each species at different regions.
# Function returns a list (each species migrated by region)
# ------------------------------------------------------------------------------

region_migrate <- function(data) {
 
  # we need dplyr for filter function
  # we need FlyingR for migration
  require(dplyr)
  require(FlyingR)
  
  # number of regions in dataset
  regions <- unique(data$Region)
    
  # regions data in a list
  region_data <- list()
  
  
  for (i in 1:length(regions)) {
    
    # filter data by region
    filter_data_region <-data %>% filter(Region == regions[i])
    
    # migrate filter data
    user_settings <- list(ipf = 0.9)
    results <- migrate(data = filter_data_region[, 3:9], method = "csw",
                          speed_control = 0, 
                          settings = user_settings,
                          min_energy_protein = 0.05)
    
    region_data[[i]] <- cbind(filter_data_region, "range" = as.vector(results$range))
    
  }
  
  # return region data as a list
  return(region_data)
  
}


```



```{r}
# ------------------------------------------------------------------------------
# migrating sylvia borin
# ------------------------------------------------------------------------------
data("garden_wablers")
wablers_migration <- region_migrate(data = garden_wablers)

```


```{r}
# ------------------------------------------------------------------------------
# migrating Lesser Whitethroats (sylvia curruca)
# ------------------------------------------------------------------------------
data("lesser_whitethroats")
whitethroats_migration <- region_migrate(data = lesser_whitethroats)

```

```{r}
# ------------------------------------------------------------------------------
# migration list to one dataframe for Garden Wablers(sylvia_borin) 
# ------------------------------------------------------------------------------
wablers_results <- data.frame()

for (i in 1:length(wablers_migration)) {
  
  wablers_results <- rbind(wablers_results, wablers_migration[[i]])
}

cat("number of rows sylvia borin:", nrow(wablers_results), sep = " ")
```

```{r}
# ------------------------------------------------------------------------------
# migration list to one dataframe for Lesser Whitethroats
# ------------------------------------------------------------------------------

whitethroats_results <- data.frame()

for (i in 1:length(whitethroats_migration)) {
  
  whitethroats_results <- rbind(whitethroats_results, whitethroats_migration[[i]])
}

cat("number of rows in sylvia curruca:", nrow(whitethroats_results), sep = " ")
```

```{r}
# ------------------------------------------------------------------------------
# order of region
# rename the regions for clarity
# ------------------------------------------------------------------------------

wablers_results$Region <- factor(wablers_results$Region,
                               levels = c("S Balt", "N Med",
                                          "E Med"), ordered = TRUE)

levels(wablers_results$Region) <- c("S Baltic", "N Medit", 
                           "E Medit")

whitethroats_results$Region <- factor(whitethroats_results$Region,
                                 levels = c("S Balt", "N Med",
                                          "E Med", "NE Afr"), ordered = TRUE)

levels(whitethroats_results$Region) <- c("S Baltic", "N Medit", 
                           "E Medit", "NE Africa")
```

```{r}
# ------------------------------------------------------------------------------
# both plots in one
# combine the data sets first
# ------------------------------------------------------------------------------

results <- rbind(wablers_results, whitethroats_results, deparse.level = 0)

# make sure it sums up
nrow(wablers_results) + nrow(whitethroats_results) == nrow(results)

```

```{r}
# ------------------------------------------------------------------------------
# new combined plot
# ------------------------------------------------------------------------------
require(ggplot2)
results_combined_plot <- ggplot(results, aes(x = Region, y = range, 
                                             colour = species)) +
  geom_boxplot() +
  #ggtitle("Sylvia curruca flight range by region") +
  theme_bw()+
  labs(y = "Range")+
  ylim(500, 4500)

results_combined_plot
```

Based on these results the two species clearly differ in their potential flight 
ranges and therefore show species-specific migration strategy, when travelling 
along the same SE flyway, as was described in the original paper based on Flight 
(Ożarowska 2015).

### Examples 2: Curlew Sandpiper
The Curlew Sandpiper is a small wader, breeding in Arctic Siberia and wintering in Africa
(Cramp and Simmons 1983). A sample of 12 individuals was observed over five days in
laboratory conditions (in total a sample of 60 observations) at the University of Gdańsk.
Initially, all these individuals had no subcutaneous fat depots (fat score equal to 0) (Meissner
2009). Birds were fed ad-libitum with mealworm larvae Tenebrio molitor. To determine the
maximum body mass gain sandpipers were weighed every day in the morning. The data set
consists of body mass, fat mass, constant wingspan (0.44 m) and wing area (0.018 m2). Body
mass and fat mass are converted from grams to kilograms. On day one, the fat fraction is
assumed to be 0.05 according to the observations of closely related species - the Red Knot
Calidris canutus (Piersma, Bruinzeel, Drent, Kersten, Van der Meer, and Wiersma 1996). On
day 2 the fat fraction increases on average to 0.23, day 3 – 0.34, day 4 – 0.40 and day 5 –
0.44. Data collected in subsequent days were used to calculate the potential flight range of the
studied individuals each day (Figure 4). On these data, the constant muscle mass migration
simulation is used. The results return a named vector, if the bird identifier is not unique, the
id suffix is added after an underscore. The results also return the speed at start and end of
migration. All these estimates may be used in the analyses of inter-individual variation of
potential flight range with increasing fat fraction in migrating birds.

```{r}
data("curlew_sandpiper")
```


```{r}
# needed data for migration columns 2:9

curlew_range <-
    migrate(
        data = curlew_sandpiper[, 2:9],
        method = "cmm",
        speed_control = 0,
        min_energy_protein = 0.05,
        settings = list(ipf = 0.9)
    )


# Addin computed range to the dataset
curlew_results <- curlew_sandpiper[, 1:2]
curlew_results$range <- as.vector(curlew_range$range)

```




## Visualize the results
```{r, message=FALSE}
# migrate curlew sandpiper
require(ggplot2)
require(dplyr)

curlew_plot <- ggplot(data = curlew_results)+
  geom_point(aes(x = day, y = range, color = name))+
  theme_bw(base_size = 10)+
  ggtitle("")+
  scale_color_discrete(name = "Individual")+
    xlab("Day")+
    ylab("Range (km)")

curlew_plot
```

### Examples 3: Various bird species data migration
Data *birds*  is a preset birds data, extracted from Flight Pennycuick(2008). Fat mass percentage
generated randomly where zero. In addition, by default muscle mass was
derived as 0.17 fraction of the all-up mass. 

## The data
   
```{r }
head(birds)
```

```{r example}
require(FlyingR)
## basic example code

## birds comes with the package
data("birds")

simulation <- migrate(data = birds,  method = "cmm", settings = list(airDensity = 0.905))

simulation$range
```

The function also returns the mechanical and chemical power during the simulation

### Range estimation based on ODE
This function estimates the range based on Pennycuick (1975) Mechanics of Flight where Breguet set of equations are used.
```{r}
## when estimating range of a single bird
birds_range <- flysim(data = birds,  settings = list(airDensity = 0.905))

birds_range$range
```




## References
Lindström Å, Piersma T (1993). “Mass changes in migrating birds: the evidence for fat
and protein storage re-examined.” Ibis, 135(1), 70–78. doi:10.1111/j.1474-919X.1993.
tb02811.x

Klaassen M (1996). “Metabolic constraints on long-distance migration in birds.” Journal of
Experimental Biology, 199(1), 57–64. doi:10.1242/jeb.199.1.57.

Kaiser A (1999). “Stopover strategies in birds: a review of methods for estimating stopover
length.” Bird Study, 46(sup1), S299–S308. doi:10.1080/00063659909477257.

Schwilch R, Grattarola A, Spina F, Jenni L (2002). “Protein loss during long-distance migra-
tory flight in passerine birds: adaptation and constraint.” Journal of Experimental Biology,
205(5), 687–695. doi:10.1242/jeb.205.5.687.

Pennycuick CJ (2008). MODELLING THE FLYING BIRD, volume 5. First edition edition.
Elsevier, New Jersey.

Lindström Å, Piersma T (1993). “Mass changes in migrating birds: the evidence for fat
and protein storage re-examined.” Ibis, 135(1), 70–78. doi:10.1111/j.1474-919X.1993.
tb02811.x.

Pearson DJ, Lack PC (1992). “Migration patterns and habitat use by passerine and
near-passerine migrant birds in eastern Africa.” Ibis, 134(s1), 89–98. doi:10.1111/j.
1474-919X.1992.tb04738.x.

Shirihai H, Gargallo G, Helbig AJ (2001). SYLVIA WARBLERS: IDENTIFICATION, TAX-
ONOMY AND PHYLOGENY OF THE GENUS SYLVIA. Christopher Helm Publishers
Ltd.

Ożarowska A (2015). “Contrasting fattening strategies in related migratory species: the
blackcap, garden warbler, common whitethroat and lesser whitethroat.” Annales Zoologici
Fennici, 52(1–2), 115–127. doi:10.5735/086.052.0210.

Bruderer B, Boldt A (2001). “Flight characteristics of birds: I. Radar measurements of
speeds.” Ibis, 143(2), 178–204. doi:10.1111/j.1474-919X.2001.tb04475.x.

Cramp S, Simmons K (1983). HANDBOOK OF THE BIRDS OF EUROPE, THE MIDDLE
EAST AND NORTH AFRICA. THE BIRDS OF THE WESTERN PALEARCTIC: 3.
WADERS TO GULLS. Oxford University Press.

Meissner W (2009). “A classification scheme for scoring subcutaneous fat depots of shore-
birds.” Journal of Field Ornithology, 80(3), 289–296. doi:10.1111/j.1557-9263.2009.
00232.x.

Piersma T, Bruinzeel L, Drent R, Kersten M, Van der Meer J, Wiersma P (1996). “Vari-
ability in basal metabolic rate of a long-distance migrant shorebird (red knot, Calidris
canutus) reflects shifts in organ sizes.” Physiological Zoology, 69(1), 191–217. doi:
10.1086/physzool.69.1.30164207

