---
title: "simulations"
author: "Brian Masinde (brima748)"
date: "2/6/2020"
output: pdf_document
---
```{r message=FALSE, warning=FALSE}
# required libraries
library(flying)
library(ggplot2)
```

# Sensitivity Analysis Simulations

## Acrocephalus scirpaceus (Eurasian Reed Wabler)
Available data on the Eurasian Reed Wabler has the following relevant 
information:

* Age of samples (All are immature)

* Fat score (0-8) 

* Samples

* Mean body mass (Kg): This is the all-up body mass

* Fat mass (kg)

* Fat fraction: Fraction of fat mass to all-up mass.

* Average flight height.

* Average wing span

* Average wing area

Note the muscle mass is not included, but can estimated as a fraction as the 
all-up mass.

So obviously we have point estimates and not the actuall data. Via sampling from 
sensible distribution with sensible parameters we could investigate the relationship
between this variables and the output (i.e the range)




```{r}
species <- rep("ACR.IRP", times = 8)
age <- rep("Immature", times = 8)
fat_score <- 0:7
mean_mass <- c(0.01127, 0.01152, 0.01152, 0.01246, 0.01246, 0.01335, 0.01335, 0.01421)
fat_mass <- c(0.00000, 0.00025, 0.00049, 0.00120, 0.00158, 0.00209, 0.00246, 0.00246)
fat_fraction <- c(0.00, 0.02, 0.04, 0.10, 0.12, 0.16, 0.18, 0.21)
wing_span_avg <- rep(0.1946, times = 8)
wing_area_avg <- rep(0.00773, times = 8)
ACR_species <- data.frame(cbind(species, age, fat_score, mean_mass, fat_mass, fat_fraction))

ACR_species
```


### Relationship between the variables and output (range)

#### Unifrom distributed measurements.

```{r}
set.seed(2020)
acr_irp_body_mass <- sort(runif(1000, min = 0.01152, max = 0.01421),
                          decreasing = FALSE)

acr_irp_fat_mass <- sort(runif(1000, min = 0.00025138, max = 0.002942),
                         decreasing = FALSE)

acr_sim_data <- data.frame(cbind("body mass" = acr_irp_body_mass, 
                                 "fat mass" = acr_irp_fat_mass))

wing.span <- sort(runif(n = 1000, min = (0.1946 - 0.00201) ,max = (0.1946 + 0.00201)))
# assuming wing area and wing span have a linear relationship, using the same scale:
wing.area <- rep(runif(n = 1000, min = 0.00773 - 0.0001 , max = 0.00773 + 0.0001))
taxon <- rep(1, 1000)
muscle.mass <- 0.09 * acr_irp_body_mass

data <- data.frame(cbind(acr_irp_body_mass, acr_irp_fat_mass, wing.span,
                         wing.area, taxon, muscle.mass))
```

```{r}
ggplot(data, aes(x = acr_irp_body_mass, y = acr_irp_fat_mass)) +
  geom_point() +
  ggtitle("Uniformly genereated data")
```

As expected there is a linear relationship between fat mass and body mass.
This is because the fat mass is a fraction of the body mass.

From research (Autumn and spring migration of the Reed Warbler Acrocephalus scirpaceus in Egypt—some interesting aspects and questions) wing span of eurasian reed wablers could have a standard deviation of 2.01mm (0.00201)

```{r}
ggplot(data, aes(x = wing.area, y = wing.span)) +
  geom_point() +
  ggtitle("Wing span vs wing area")
```


```{r}
ggplot(data, aes(x = wing.span, y = acr_irp_body_mass)) +
  geom_point() +
  ggtitle("Wing span vs wing area")
```

```{r}
ggplot(data, aes(x = wing.span, y = acr_irp_fat_mass)) +
  geom_point() +
  ggtitle("Wing span vs wing area")
```

```{r message=FALSE, paged.print=FALSE}
test_results <- migrate(data = data, method = "cmm", speed_control = "constant_speed")

ggplot(data, aes(x = acr_irp_body_mass, y = test_results$range)) +
  geom_point() +
  ggtitle("Uniformly genereated data")
```


```{r}
#plot(density(test_results$range), main = "Range density, uniform body mass and fat mass")
set.seed(2020)
ggplot()+
  geom_density(aes(x = test_results$range), fill = "#66CDAA")+
  geom_density(aes(x = runif(n = 1000, min = min(test_results$range), max = max(test_results$range)))
               , fill = "#B0E0E6", alpha = 0.6) +
  ggtitle(label = "Comparison of the range estimates to a unfirom distribution") +
  xlab("range (Km)") +
  ylab("Density")
```

The range results follow the same distribution as that of the body mass/ fat mass.
The algorithm "constant muscle mass" depends on fat mass.


###  Gaussian body mass and fat  mass.
```{r}
# normally distributed data
set.seed(2020)
acr_irp_body_mass <- rnorm(1000, mean = 0.01175, sd = 0.001)

acr_irp_fat_frac <- rnorm(1000, mean = 0.04170213, sd = 0.00001)

acr_irp_fat_mass <- acr_irp_body_mass * acr_irp_fat_frac

acr_sim_data <- data.frame(cbind("body mass" = acr_irp_body_mass, 
                                 "fat mass" = acr_irp_fat_mass))

data <- data.frame(cbind(acr_irp_body_mass, acr_irp_fat_mass, wing.span,
                         wing.area, taxon, muscle.mass))

ggplot(acr_sim_data, aes(x = fat.mass, y = body.mass)) +
  geom_point() +
  ggtitle("Normall distributed genereated data")
```


```{r}
test_results <- migrate(data = data, method = "cmm", speed_control = "constant_speed")

ggplot(acr_sim_data, aes(x = body.mass, y = test_results$range)) +
  #geom_line() +
  geom_point()+
  ggtitle("Range vs body mass")
```



```{r message=FALSE}
test_results <- migrate(data = data, method = "cmm", speed_control = "constant_speed")

ggplot(acr_sim_data, aes(x = fat.mass, y = test_results$range)) +
  geom_point() +
  ggtitle("Range vs body mass")
```

```{r}
plot(density(test_results$range), main = "Range density, gaussian body mass and fat mass")
```



The data suggests that the bird samples could come from a mixture of distributions
within the immature demographic.

```{r}
# sample body mass from a gaussian mixture.
set.seed(2020)
avg <- c(827,2160, 180, 97,  94,  51,  25)/3434
body_mass_mean <- c(0.01152, 0.01175, 0.01246, 0.01246, 0.01335, 0.01373, 0.01421)
fat_mass_mean <- c(0.00025, 0.00049, 0.00120, 0.00158, 0.00158, 0.00246, 0.00246)

probs <- c(0.241,0.629, 0.0524, 0.0282, 0.0274,  0.0149, 0.00728)
probs2 <- fat_mass_mean/sum(fat_mass_mean)
body_mass <- c()
fat_mass <- c()
for (i in 1:1000) {
  # choose which component to sample from
  comp <- which(rmultinom(1, size = 1, prob = probs) == 1)
  comp2 <- which(rmultinom(1, size = 1, prob = probs2) == 1)
  
  # sanple from gaussin with 10% variation
  body_mass[i] <- rnorm(1, mean = body_mass_mean[comp], sd = 0.1 *body_mass_mean[comp])
  fat_mass[i] <- rnorm(1, fat_mass_mean[comp2], sd = 0.1*fat_mass_mean[comp2])
}

```

```{r}
ggplot()+
  geom_density(aes(x = body_mass), fill = "#66CDAA")+
  #geom_density(aes(x = runif(n = 1000, min = min(test_results$range), max = max(test_results$range)))
  #             , fill = "#B0E0E6", alpha = 0.6) +
  ggtitle(label = "Distribution of body mass") +
  xlab("Body mass") +
  ylab("Density")
```

```{r}
ggplot()+
  geom_density(aes(x = fat_mass), fill = "#66CDAA")+
  #geom_density(aes(x = runif(n = 1000, min = min(test_results$range), max = max(test_results$range)))
  #             , fill = "#B0E0E6", alpha = 0.6) +
  ggtitle(label = "Distribution of fat mass") +
  xlab("fat mass") +
  ylab("Density")
```


```{r}
# relationship between fat mass and muscle mass
ggplot() + 
  geom_point(aes(x = fat_mass, y = body_mass), color = "#C6103F")+
  ggtitle("Relationship between fat and body mass")+
  xlab("fat mass")+
  ylab("body mass")
```

```{r}
# wing span from normal distribution
set.seed(2020)
wing_span <- rnorm(1000, mean = 0.1946, sd = 0.01 * 0.1946)
wing_area <- rnorm(1000, mean = 0.00773, sd = 0.01 * 0.00773)
muscle_mass <- 0.15 * body_mass
taxon <- rep(1, time = 1000)
norm_gen_data <- data.frame(cbind(body_mass, fat_mass, wing_span, wing_area,
                                  muscle_mass, taxon))

results <- migrate(data = norm_gen_data, method = "cmm", 
                   speed_control = "constant_speed")
```

```{r}
# relationship between fat mass and range
ggplot()+
  geom_point(aes(x = fat_mass, y = results$range))+
  ggtitle("Relationship between fat mass and range")+
  xlab("fat mass") +
  ylab("range")
```

```{r}
ggplot()+
  geom_point(aes(x = muscle_mass, y = results$range))+
  ggtitle("Relationship between muscle mass and range")+
  xlab("fat mass") +
  ylab("range")
```

```{r}
ggplot()+
  geom_point(aes(x = body_mass, y = results$range))+
  ggtitle("Relationship between body mass and muscle mass")+
  xlab("body mass") +
  ylab("range")
```


```{r}
ggplot()+
  geom_density(aes(x = results$range))+
  ggtitle("Distribution of range")+
  xlab("range") +
  ylab("density")
```


### Uncertainty in mechanical conversion efficiency.

Default value in flight is 0.23

```{r}
# uniform prior 
set.seed(2020)
mce <- runif(100, min = 0.09, max = 0.23)

boxplot(mce, main = "Variation in mechanical conversion efficiency")
```


Fat mass and muscle mass as constants
```{r message=FALSE, warning=FALSE}
body.mass <- 0.01175
fat.mass <- 0.00049
muscle.mass <- body.mass * 0.17
wing.span <- 0.1946
wing.area <- 0.00773
taxon <- 1

data <- data.frame(cbind(body.mass, fat.mass, muscle.mass, wing.span, wing.area, taxon))

res <- list()
for (i in 1:length(mce)) {
  res[[i]] <- migrate(data = data, method = "cmm", speed_control = "constant_speed",
                      settings = list(mce = mce[i]) )
}
```

```{r}
ranges <- c()

for (i in 1:length(mce)) {
  ranges[i] <- res[[i]][[1]]
}

ggplot() +
  geom_line(aes(x = mce, y = ranges)) +
  ggtitle("Range vs body mass")
```

#### Larger variation in mce
```{r}
# uniform prior 
set.seed(2020)
mce <- rnorm(100,mean = 0.23, sd = 0.02)

boxplot(mce, main = "Variation in mechanical conversion efficiency")
```

```{r message=FALSE, warning=FALSE}
body.mass <- 0.01175
fat.mass <- 0.00049
muscle.mass <- body.mass * 0.17
wing.span <- 0.1946
wing.area <- 0.00773
taxon <- 1

data <- data.frame(cbind(body.mass, fat.mass, muscle.mass, wing.span, wing.area, taxon))

res <- list()
for (i in 1:length(mce)) {
  res[[i]] <- migrate(data = data, method = "cmm", speed_control = "constant_speed",
                      settings = list(mce = mce[i]) )
}

ranges <- c()

for (i in 1:length(mce)) {
  ranges[i] <- res[[i]][[1]]
}

ggplot() +
  geom_line(aes(x = mce, y = ranges)) +
  ggtitle("Range vs mechanical conversion efficiency")
```
