---
title: "Uncertainty Analysis Garden Wablers"
author: "Brian Masinde (brima748)"
date: "3/2/2020"
output: pdf_document
---
```{r}
library(ggplot2)
library(readr)
library(dplyr)
library(xtable)
library(gridExtra)
```

# The DATA
```{r}
library(readr)

sylvia_borin_data <- read_delim("sylvia_borin_data.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)

# remove x4 (extra column not needed)

sylvia_borin_data$X4 <- NULL
View(sylvia_borin_data)
```

## Summary table
```{r}
sb_data <- sylvia_borin_data
sb_data$Fat <- as.factor(sb_data$Fat)
summary_sb_data <- summary(sylvia_borin_data)

# save this summary table to latex
print.xtable(summary_sb_data, ttype = "latex", file = "summary_sb.tex")
```

## Distribution of the body mass
```{r}
plot(density(sylvia_borin_data$Mass/1000, na.rm = TRUE), 
     main = "sylvia borin", xlab = "body mass in Kg")
```

```{r}
bodymass_distribution <- ggplot(data = sylvia_borin_data)+
  geom_density(aes(x = Mass), fill = "#66CDAA") +
  xlab("Body mass (grams)") +
  theme_minimal()+
  ggtitle("Body mass distribution")
  
ggsave(filename = "bodymass_distribution.png",
       device = png(), dpi = 300, plot = last_plot(),
       width = 7, height = 7, units = "cm")
```


```{r}
summary(sylvia_borin_data$Mass)
```

From the density plot, the shape exclude uniform distribution. The values have
to be positive, exclude normal distribution. However, the shape is just about gaussian.
Distributions that mimick gaussian distribution but non-negative i.e Gamma distribution,
and Cauchy distribution.

According to \textit{ottosson2005garden} the garden wabler is a classic subject
for the study of Palaearctic–African bird migration strategies. They state that 
the true depature mass of the bird might be between 25 and 30 grams.

```{r}
mean_sylvia_borin <- mean(sylvia_borin_data$Mass, na.rm = TRUE)

mean_sylvia_borin
```


```{r}
moment_2 <- mean(sylvia_borin_data$Mass^2, na.rm = TRUE)

moment_2
```

```{r}
# alpha
alpha <-  mean_sylvia_borin^2/(mean(moment_2 - mean_sylvia_borin^2, na.rm = TRUE))

alpha
```

```{r}
# beta
beta <- mean_sylvia_borin/(mean(moment_2 - mean_sylvia_borin^2, na.rm = TRUE))

beta
```


```{r}
sim_gamma <- rgamma(2000, shape = alpha, rate = beta)
```


```{r}
plot(density(sylvia_borin_data$Mass, na.rm = TRUE))
lines(density(sim_gamma))
```


## Fat mass.
\textit{ottosson2005garden} say that 45 -50% of body mass may be fuel reserves
alone (Autumn). Body mass is positively correlated with fat score (implies there
might be a correlation with fat mass also).

```{r}
hist(sylvia_borin_data$Fat)
#lines(density(sylvia_borin_data$Mass))
```

```{r}
# ggplot histogram of fat score.
fat_score <- ggplot(sb_data, aes(x = Fat, color = Fat, fill = Fat)) +
  geom_histogram(stat = "count")+
  theme_minimal()+
  xlab("Fat score")+
  ggtitle("Fat score distribution")
  
ggsave("fat_score.png", device = png(), dpi = 300, plot = last_plot(),
       width = 7, height = 7, units = "cm")
```

```{r}
wl_distribution <- ggplot(data = sylvia_borin_data)+
  geom_density(aes(x = Wing), fill = "#66CDAA") +
  xlab("Wing length (mm)") +
  theme_minimal()+
  ggtitle("Wing length distribution")
  
# ggsave(filename = "bodymass_distribution.png",
#        device = png(), dpi = 300, plot = last_plot(),
#        width = 7, height = 7, units = "cm")
grid_plt <- grid.arrange(bodymass_distribution, fat_score, wl_distribution,
                         ncol = 2, nrow = 2)
ggsave("grid_plt.png", device = png(), dpi = 300, plot = grid_plt,
       width = 14, height = 14, units = "cm")
```



## Wing span

# plot of the wing length
```{r}
plot(density(sylvia_borin_data$Wing, na.rm = TRUE), 
     main = "sylvia borin", xlab = "wing length")
```

```{r}
# transform by 2.9
ws <- sylvia_borin_data$Wing  * 2.9

```


## Wing area
Estimaing wing area from wing length usign linear regression

```{r}

wing_area_sim <- -30.167  + 1.616 * sylvia_borin_data$Wing

plot(density(wing_area_sim, na.rm = TRUE))
```

```{r}

```


## muscle mass

## Flying Altitude


## 




