---
title: "Uncertainty Analysis"
author: "Brian Masinde (brima748)"
date: "2/19/2020"
output: 
  pdf_document:
  citation_package: biblatex
bibliography: references.bib
---
```{r}
library(ggplot2)
library(flying)
library(corrplot)
library(mgcv)
```

# Uncertainty Analysis in Preset birds.

## Anas Crecca (Eurasian Teal)

### Body Mass (Kg)
@fox1992seasonal measured the Teal at different times of the year. They present a table
of samples between adults, immature, males and females. In summary,
among juvenile males; sample of 3577 had the body mass of $303.6 \pm 0.6$ while
adult males (999) have the distribution of $320.4 \pm 1.1$, juvinile females
(2295) $280.4 \pm 0.7$ and adult females $290.6 \pm 1.4$.

This makes it possible to sample from a gaussian mixture distribution.

```{r}
set.seed(2020)
anas_crecca_BM_sample <- c()

# sample order (juvenile male, adult male, juvenile female, adult female)
samples <- c(3577, 999, 2295, 498)
# probabilities
probs <- samples/sum(samples)
# standard deviations
sd <- c(0.6, 1.1,0.7, 1.4)
# means of body mass
body_mass_mean <- c(303.6, 320.4, 280.4, 290.6)

for (i in 1:1000) {
  # choose which component to sample from
  comp <- which(rmultinom(1, size = 1, prob = probs) == 1)
  
  # sanple from gaussin with 10% variation
  anas_crecca_BM_sample[i] <- rnorm(1, mean = body_mass_mean[comp], sd = sd)
}
```


```{r}
ggplot()+
  geom_density(aes(x = anas_crecca_BM_sample), fill = "grey")+
  xlim(265, 330)+
  theme_classic()+
  ggtitle("Mixture sample of Anas Crecca Body Mass (grams)")
```

### Wing Span.
According to https://www.hbw.com/species/common-teal-anas-crecca (Handbook of the 
birds of the world) the wing span of the birds could vary from 54 - 62 cm. Sample
from a uniform distribution.

```{r}
set.seed(2020)
anas_crecca_WS_sample <- runif(1000, min = 54, max = 62)
```

```{r}
ggplot()+
  geom_density(aes(x = anas_crecca_WS_sample), fill = "grey")+
  theme_classic()+
  ggtitle("Wing Span distirbution Anas crecca")
```

### Uncertainty in Wing Area.
Wing area has some sort of linear relationship with the wing span. Can we use 
linear regrssion to predict wing area?

```{r}
WAWS <- ggplot()+
  geom_point(aes(x = log(birds$Wing.area), y = log(birds$Wing.span)), color = "#145A32")+
  xlab("wing area")+
  ylab("wing span")+
  ggtitle("Wing area vs Wing span (Log Transformed)")+
  theme_bw()
WAWS
```

```{r}
# fitting a linear regression
lm_wswa <- lm(Wing.area ~ Wing.span, data = birds)
```

```{r}
ggplot()+
  geom_density(aes(x = lm_wswa$residuals))+
  geom_density(aes(x = rnorm(1000, mean = 5.138344e-18, sd = 0.1513841)))+
  theme_classic()+
  xlim(-0.5, 0.5)
```

```{r}
plot(lm_wswa)
```

Not so nice prediction. But we can still use it to predict

```{r}
anas_crecca_WingS_sample <- data.frame(cbind("Wing.span" = anas_crecca_WS_sample))
predicted_WA <- predict.lm(object = lm_wswa, anas_crecca_WingS_sample)
```

```{r}
ggplot()+
  geom_point(aes(x = anas_crecca_WS_sample, y = predicted_WA))+
  ggtitle("Predicted wing area vs wing span")
```

### Uncertainty in fat mass
Suppose that bird could vary in obesity. Usually fat mass is not directly observable.
Fat scores are widely used. We dont have fat score relationship betweeen the teals.
Way out we could sample the fat fraction form a uniform distribution. This is better
because a big bird does not necessariy have to be obesse. During migration,
birds can have between $50-60\%$  of body mass as fat [@guglielmo2018obese].

```{r}
# simulating from a uniform distribution.
set.seed(2020)
anas_crecca_fat_fraction <- runif(1000, min = 0.2, max = 0.6)
```

### Uncertainty in muscle mass.
Again muscle mass is not directly observable but it is a fraction of the body mass.
Couldnt find a standard for this measurement. So we sample from reasonable uniform
range. 

```{r}
# most birds in flight have a musscle fraction of 0.17
set.seed(2020)
anas_crecca_muscle_frction <- rnorm(1000, mean = 0.17, sd = 0.01)
```

```{r}
ggplot()+
  geom_density(aes(x= anas_crecca_muscle_frction), fill = "grey")+
  xlim(0.13, 0.21)+
  theme_classic()
```

```{r}
# aggregate data on anas crecca
anas_crecca_bodymass <- round((anas_crecca_BM_sample*0.001), 5)
anas_crecca_wingspan <- round((anas_crecca_WS_sample) * 0.01, 5)
anas_crecca_wingarea <- round(predicted_WA * 0.01,5)
anas_crecca_musclemass <- round((anas_crecca_BM_sample * anas_crecca_muscle_frction)*0.001,5)
anas_crecca_fatmass <- round((anas_crecca_BM_sample * anas_crecca_fat_fraction)*0.001,5)
species <- rep("Anas Crecca", 1000)
taxon <-  rep(2, 1000)

anas_crecca <- data.frame(
  cbind(species,
        anas_crecca_bodymass,
        anas_crecca_wingspan,
        anas_crecca_wingarea,
       taxon,
       anas_crecca_musclemass,
        anas_crecca_fatmass
  ),
  stringsAsFactors = FALSE
)

colnames(anas_crecca) <- c("species","body.mass","wing.span","wing.area","taxon","muscle.mass",
                           "fat.mass")
anas_crecca$species <- as.factor(anas_crecca$species)
anas_crecca$taxon <- as.factor(anas_crecca$taxon)
anas_crecca$body.mass <- as.numeric(anas_crecca$body.mass)
anas_crecca$fat.mass <- as.numeric(anas_crecca$fat.mass)
anas_crecca$wing.span <- as.numeric(anas_crecca$wing.span)
anas_crecca$wing.area <- as.numeric(anas_crecca$wing.area)
anas_crecca$muscle.mass <- as.numeric(anas_crecca$muscle.mass)

```



### Simulate range
```{r}
anas_crecca_range <- migrate(data = anas_crecca, method = "csw", speed_control = "constant_speed")
```

```{r}
rho <- cor(x = data.frame(cbind(anas_crecca_bodymass, anas_crecca_wingspan, anas_crecca_wingarea,anas_crecca_musclemass, anas_crecca_fatmass, anas_crecca_range$range)))
corrplot(rho)
```

```{r}
ggplot()+
  geom_density(aes(log(anas_crecca_range$range)), fill = "grey")+
  xlim(6, 8.9)
```

```{r}
ggplot()+
  geom_point(aes(x = anas_crecca$fat.mass, y = anas_crecca_range$range))+
  theme_bw()
```

## Calidris Tenuirostris
According to https://www.hbw.com/species/red-knot-calidris-canutus, great knots
can have body mass distribution anywhere between $85-220g$,but @penny03 observed birds
with mean body mass o f 233g.
Reasonable uncertainty uniform distribution
```{r}
set.seed(2020)
red_knot_BM_sample <- runif(1000, min = 0.1, max = 0.233)
# wing span
red_knot_WS_sample <- runif(1000, min = 0.45, max = 0.6)
# wing area
red_knot_WA_sample <- rnorm(1000, mean = 0.0397, sd = 0.00183)

red_knot_FF_sample <- runif(1000, min = 0.2, max = 0.6)

red_knot_MF_sample <- rnorm(1000, mean = 0.144, sd = 0.01)
```

@penny03 give mean and sd of wing span and wing area.

## Garden Wabler.

### Uncertainty Analysis
According to hbw.com garden wabler can range from 16 to 22 grams and upto 37 grams
prior to migration. Assuming normal distribution with mean around 22 grams with 
variance.
```{r}
set.seed(2020)
# body mass
garden_wabler_BM_sample <- round(rnorm(1000, mean = 0.025, sd = 0.001), 5)

# Fat fraction using same strategy 50 -60 %
garden_wabler_FF_sample <- round(runif(1000, min = 0.2, max = 0.6), 5)

# muscle mass
garden_wabler_MF_sample <- round(rnorm(1000, mean = 0.144, sd = 0.01), 5)

# pertubate the wing span slightly
garden_wabler_WS_sample <- round(rnorm(1000, mean = 0.24, sd = 0.001), 5)

# pertubate wing area
garden_wabler_WA_sample <- round(rnorm(1000, mean = 0.011, sd = 0.001), 5)

species <- rep("Sylvia borin", 1000)

taxon <-  rep(1, 1000)
```


```{r}
garden_wabler_fat_mass <- garden_wabler_BM_sample * garden_wabler_FF_sample
garden_wabler_muscle_mass <-garden_wabler_BM_sample*garden_wabler_MF_sample

garden_wabler_data <- data.frame(
  cbind(
    garden_wabler_BM_sample,
    garden_wabler_fat_mass,
    garden_wabler_muscle_mass,
    garden_wabler_WS_sample,
    garden_wabler_WA_sample,
    species,
    taxon,deparse.level = 0
  ),
  stringsAsFactors = FALSE
)

colnames(garden_wabler_data) <- c("body.mass","fat.mass","muscle.mass", "wing.span","wing.area",
                           "species","taxon")
garden_wabler_data$species <- as.factor(garden_wabler_data$species)
garden_wabler_data$taxon <- as.factor(garden_wabler_data$taxon)
garden_wabler_data$body.mass <- as.numeric(garden_wabler_data$body.mass)
garden_wabler_data$fat.mass <- as.numeric(garden_wabler_data$fat.mass)
garden_wabler_data$wing.span <- as.numeric(garden_wabler_data$wing.span)
garden_wabler_data$wing.area <- as.numeric(garden_wabler_data$wing.area)
garden_wabler_data$muscle.mass <- as.numeric(garden_wabler_data$muscle.mass)
```

```{r}
# migrate garden wabler
garden_wabler_range <- migrate(data = garden_wabler_data, method = "csw", speed_control = "constant_speed")
```
```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range))+
  theme_bw()
```


```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$body.mass, y = garden_wabler_range$range))+
  theme_bw()
```


```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$muscle.mass, y = garden_wabler_range$range))+
  theme_bw()
```

```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$wing.span, y = garden_wabler_range$range))+
  theme_bw()
```


```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$wing.area, y = garden_wabler_range$range))+
  theme_bw()
```

### Sensitivity Analysis

#### Variance Based Method (SOBOL):

```{r}
# Dont run
# expected value at each interval 
expected_value <- function(y,x, r) {
  # x anyone of the inputs
  # y estimated range
  # number of values to take
  
  if(length(x) != length(y)) {
    stop("lenght not equal")
  }
  n <- length(y)
  chunks <- n/r
  #min_x <- min(x)
  #max_x <- max(x)
  #last_x <- min_x
  #z <-  1# counter
  predicted <- c() # mean_value
  for (i in 1:chunks) {
    predicted[i] <- mean(y[i:r*i])
  }
  return(predicted)
}
```

```{r}
preds <- expected_value(garden_wabler_range$range, garden_wabler_data$wing.area, 5)
```

```{r}
try_model <- lm(garden_wabler_range$range ~ wing.area + 
                      wing.area^2,
                    data = garden_wabler_data)

preds <- predict(try_model)
```

```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$wing.area, y = garden_wabler_range$range))+
  geom_line(aes(x = garden_wabler_data$wing.area, y = preds), color  = "red")+
  geom_smooth(aes(x = garden_wabler_data$wing.area, y = garden_wabler_range$range))+
  theme_bw()
```

```{r}
# sobol index of wing area

S_wing_area <- var(preds)/var(garden_wabler_range$range)

S_wing_area
```

```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range))+
  geom_smooth(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range),
              span = 0.8)+
  theme_bw()
```

```{r}
fat_range_gam <- lm(formula = garden_wabler_range$range ~ fat.mass + 
                      fat.mass^2 + fat.mass^3 + fat.mass^4,
                    data = garden_wabler_data)
range_pred_fat <- predict(fat_range_gam)
```

```{r}
ggplot()+
  geom_point(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range))+
  geom_line(aes(x = garden_wabler_data$fat.mass, y = range_pred_fat), color  = "green")+
  geom_smooth(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range),
              span = 0.8)+
  theme_bw()
```

```{r}
gam_model <- gam(formula = garden_wabler_range$range ~ s(fat.mass, bs = 'cs'),
              data = garden_wabler_data)

pred3 <- predict(gam_model)
ggplot()+
  geom_point(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range))+
  geom_line(aes(x = garden_wabler_data$fat.mass, y = range_pred_fat), color  = "green")+
  geom_line(aes(x = garden_wabler_data$fat.mass, y =pred3), color  = "red")+
  #geom_smooth(aes(x = garden_wabler_data$fat.mass, y = garden_wabler_range$range),
   #           span = 0.8)+
  theme_bw()
```

```{r}
# sobol index of wing area
# GAM
S_fat_mass <- var(pred3)/var(garden_wabler_range$range)

S_fat_mass
```

```{r}
rho <- cor(x = data.frame(cbind(garden_wabler_data[, -c(6,7)], garden_wabler_range$range)))
corrplot(rho)
```

#### Quasi random sampling.

```{r}
set.seed(2020)
# creating a quasi monte carlo random sampling function.
dists <- list(
  bm = list(a = rnorm, mean = 0.028, sd = 0.001),
  ff = list(a = runif, min = 0.5, max = 0.6),
  mm = list(a = rnorm, mean = 0.17, sd = 0.0004),
  wa = list(a = rnorm, mean = 0.011, sd = 0.0001),
  ws = list(a = rnorm, mean = 0.24, sd = 0.0005)
)

quasi_mc <- function(N, dists) {
  d <- length(dists)
  # N2d matrix
  N2d <- matrix(data = NA, nrow = N, ncol = 2*d)
  
  for (i in 1:nrow(N2d)) {
    z <- 1
    for (j in 1:ncol(N2d)) {
      
      if (j <= d) {
        N2d[i, j] <- round(dists[[j]]$a(1, dists[[j]][[2]], dists[[j]][[3]]), 5)
        #print(i)
        #print(j)
        #print(dists[[j]]$a)
      } else {
        #z <- j %/% 2 # quotient
        N2d[i, j] <- round(dists[[z]]$a(1, dists[[z]][[2]], dists[[z]][[3]]), 5)
        z <- z+1
      }
    }
  }
  
  # replicating variable names
  cnames <- c("body.mass", "fat.frac", "muscle.frac", "wing.area", "wing.span")

  cnames <- rep(cnames, 2)

  colnames(N2d) <- cnames
  
  # first d columns as A
  A <- N2d[ ,1:d]

  # remains d columns as B
  B <- N2d[ ,(d+1):(d*2)]
  
  # create Ab_i
  AB_sub_i <- list()

  for (i in 1:d) {
    cnme <- colnames(A)[i]
    AB_sub_i[[i]] <- as.matrix(cbind(A[, i], B[, -i]))
    colnames(AB_sub_i[[i]])[1] <- cnme
  }

  # fractions to masses
  for (i in 1:length(AB_sub_i)) {
    # fat mass
    AB_sub_i[[i]] <- as.matrix(cbind(AB_sub_i[[i]],
                                     "fat.mass" = AB_sub_i[[i]][, "fat.frac"] *
                                       AB_sub_i[[i]][, "body.mass"]))
    # muscle mass
    AB_sub_i[[i]] <- as.matrix(cbind(AB_sub_i[[i]],
                                     "muscle.mass" = AB_sub_i[[i]][, "muscle.frac"] *
                                       AB_sub_i[[i]][, "body.mass"]))
    
    # remove fractions
    
    AB_sub_i[[i]] <-
      AB_sub_i[[i]][, !colnames(AB_sub_i[[i]]) %in% c("fat.frac", "muscle.frac")]
    
  }

  A2 <- as.matrix(cbind(A, "fat.mass" = round(A[, "fat.frac"] * A[, "body.mass"], 5)))

  A2 <- as.matrix(cbind(A2, "muscle.mass" = round(A[, "muscle.frac"] * A[, "body.mass"],5)))

  # remove fat.frac and fat.mass
  A2 <-  A2[, !colnames(A2) %in% c("fat.frac", "muscle.frac")]

  # convert fractions to mass
  B2 <- as.matrix(cbind(B, "fat.mass" = B[, "fat.frac"] * B[, "body.mass"]))
  B2 <- as.matrix(cbind(B2, "muscle.mass" = B[, "muscle.frac"] * B[, "body.mass"]))

  B2 <- B2[, !colnames(B2) %in% c("fat.frac", "muscle.frac")]

  
  return(list("N2d" = N2d, "A" = A2, "B" = B2, "AB_sub_i" = AB_sub_i))
} 

N <- 300
trial <- quasi_mc(N, dists)
```


```{r}
# extracting matrices
A <- trial$A

B <- trial$B

AB_sub_1 <-  trial$AB_sub_i[[1]]

AB_sub_2 <-  trial$AB_sub_i[[2]]

AB_sub_3 <-  trial$AB_sub_i[[3]]

AB_sub_4 <-  trial$AB_sub_i[[4]]

AB_sub_5 <-  trial$AB_sub_i[[5]]

# adding species and taxon
species <- rep("Sylvia borin", 10)

taxon <-  rep(1, 10)

A <- data.frame(cbind(A, "species" = species, "taxon" = taxon))
A[ , c(1,2,3,4,5)] <- apply(A[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
A[ , !c(1,2,3,4,5)] <- apply(A[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))

B <- data.frame(cbind(B, "species" = species, "taxon" = taxon))
B[ , c(1,2,3,4,5)] <- apply(B[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
B[ , !c(1,2,3,4,5)] <- apply(B[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))


AB_sub_1 <- data.frame(cbind(AB_sub_1, "species" = species, "taxon" = taxon))
AB_sub_1[ , c(1,2,3,4,5)] <- apply(AB_sub_1[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
AB_sub_1[ , !c(1,2,3,4,5)] <- apply(AB_sub_1[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))

AB_sub_2 <- data.frame(cbind(AB_sub_2, "species" = species, "taxon" = taxon))
AB_sub_2[ , c(1,2,3,4,5)] <- apply(AB_sub_2[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
AB_sub_2[ , !c(1,2,3,4,5)] <- apply(AB_sub_2[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))

AB_sub_3 <- data.frame(cbind(AB_sub_3, "species" = species, "taxon" = taxon))
AB_sub_3[ , c(1,2,3,4,5)] <- apply(AB_sub_3[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
AB_sub_3[ , !c(1,2,3,4,5)] <- apply(AB_sub_3[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))

AB_sub_4 <- data.frame(cbind(AB_sub_4, "species" = species, "taxon" = taxon))
AB_sub_4[ , c(1,2,3,4,5)] <- apply(AB_sub_4[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
AB_sub_4[ , !c(1,2,3,4,5)] <- apply(AB_sub_4[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))

AB_sub_5 <- data.frame(cbind(AB_sub_5, "species" = species, "taxon" = taxon))
AB_sub_5[ , c(1,2,3,4,5)] <- apply(AB_sub_5[ , c(1,2,3,4,5)], 2, function(x) as.numeric(as.character(x)))
AB_sub_5[ , !c(1,2,3,4,5)] <- apply(AB_sub_5[ ,!c(1,2,3,4,5)], 2, function(x) as.factor(x))

```

```{r}
range_A <- migrate(data = A, method = "csw", speed_control = "constant_speed")
range_B <- migrate(data = B, method = "csw", speed_control = "constant_speed")
range_AB_sub_1 <- migrate(data = AB_sub_1, method = "csw", speed_control = "constant_speed")
range_AB_sub_2 <- migrate(data = AB_sub_2, method = "csw", speed_control = "constant_speed")
range_AB_sub_3 <- migrate(data = AB_sub_3, method = "csw", speed_control = "constant_speed")
range_AB_sub_4 <- migrate(data = AB_sub_4, method = "csw", speed_control = "constant_speed")
range_AB_sub_5 <- migrate(data = AB_sub_5, method = "csw", speed_control = "constant_speed")
```

### First order sensitivity indices
```{r}
# approximate sobol indices

sb_index <- log(range_B$range)*(log(range_AB_sub_1$range) - log(range_A$range))

mean(sb_index)
```

```{r}
# Total effect index
1/(2*N)* (sum((log(range_A$range) - log(range_AB_sub_2$range))^2))
```


```{r}
# numerical estimation of first order indices

first_order_index <- function(rangeA, rangeB, range_AB_sub_i, d) {
  # store 
  sb_index <- vector(length = d)
  
  for (i in 1:length(range_AB_sub_i)) {
    
  }
}
```



###
```{r}

```


```{r}
ggplot()+
  geom_point(aes(x = A$body.mass, range_A$range))
  geom_smooth(aes(x = A$body.mass, range_A$range))
```

```{r}
ggplot()+
  geom_point(aes(x = A$fat.mass, range_A$range))+
  geom_smooth(aes(x = A$fat.mass, range_A$range))
```

```{r}
ggplot()+
  geom_point(aes(x = A$muscle.mass, range_A$range))+
  geom_smooth(aes(x = A$muscle.mass, range_A$range))
```

```{r}
ggplot()+
  geom_point(aes(x = c(A$fat.mass, B$fat.mass, AB_sub_1$fat.mass, 
                           AB_sub_2$fat.mass, AB_sub_3$fat.mass,
                           AB_sub_4$fat.mass, AB_sub_5$fat.mass),
                 y = c(range_A$range, range_B$range, range_AB_sub_1$range,
                           range_AB_sub_2$range, range_AB_sub_3$range,
                           range_AB_sub_4$range, range_AB_sub_5$range)))
```


## Preset birds
```{r}
preset_migration <- migrate(data = birds, method = "csw", speed_control = "constant_speed")
```
```{r}
ggplot()+
  geom_point(aes(x = birds$Fat.mass, y = preset_migration$range))+
  geom_smooth(aes(x = birds$Fat.mass, y = preset_migration$range),
              span = 0.8)+
  theme_bw()
```

```{r}
ggplot()+
  geom_point(aes(x = birds$Muscle.mass, y = preset_migration$range))+
  geom_smooth(aes(x = birds$Muscle.mass, y = preset_migration$range),
              span = 0.8)+
  theme_bw()
```

```{r}
ggplot()+
  geom_point(aes(x = birds$Wing.span, y = preset_migration$range))+
  geom_smooth(aes(x = birds$Wing.span, y = preset_migration$range),
              span = 0.8)+
  theme_bw()
```

```{r}
ggplot()+
  geom_point(aes(x = birds$Wing.area, y = preset_migration$range))+
  geom_smooth(aes(x = birds$Wing.area, y = preset_migration$range),
              span = 0.8)+
  theme_bw()
```

```{r}
ggplot()+
  geom_point(aes(x = birds$Wing.area, y = log(preset_migration$range)))+
  geom_smooth(aes(x = birds$Wing.area, y = log(preset_migration$range)),
              span = 0.8)+
  theme_bw()
```


# References
